Parsing Postfix log files

### 2008-02-20 doc/logparser.tex
Documented DELETE action and associated complication.

\newcommand{\numberOFrules}[0]{%
    142%
}

@@

@@

    \item [helo] The hostname used in the HELO command.  The HELO
        occasionally changes during a connection, presumably because spam
        or virus senders think it's a good idea.  By default Postfix only
        logs the HELO when it rejects an \SMTP{} command, but it is quite
        easy to rectify this:

@@

        Although \texttt{smtpd\_helo\_restrictions} seems like the natural
        place to log the HELO hostname, there won't be a queueid associated
        with the mail for the first recipient, so that log line cannot be
        associated with the correct mail.  There is guaranteed to be
        queueid when the DATA command has been reached, and thus it will be
        logged by any restrictions taking effect in
        \texttt{smtpd\_data\_restrictions}.  There is no difficulty in
        specifying a HELO-based restriction in
        \texttt{smtpd\_data\_restrictions}, Postfix will perform the check
        correctly.

@@

@@

>>>>

\newcommand{\numberOFrules}[0]{%
    146%
}

@@

XXX COVERAGE\@.

@@

    \item [helo] The hostname used in the HELO command.  The HELO
        occasionally changes during a connection, presumably because spam
        or virus senders think it's a good idea.  By default Postfix only
        logs the HELO hostname when it rejects an \SMTP{} command, but it
        is quite easy to rectify this:

@@

        Although \texttt{smtpd\_helo\_restrictions} seems like the natural
        place to log the HELO hostname, there won't be a queueid associated
        with the mail for the first recipient, so that log line cannot be
        associated with the correct mail.  There is guaranteed to be a
        queueid when the DATA command has been reached, and thus it will be
        logged by any restrictions taking effect in
        \texttt{smtpd\_data\_restrictions}.  There is no difficulty in
        specifying a HELO-based restriction in
        \texttt{smtpd\_data\_restrictions}, Postfix will perform the check
        correctly.

        Logging the HELO hostname in this fashion also prevents the
        complication described in \sectionref{Mail deleted before delivery
        is attempted} from occurring, because when the HELO hostname is
        logged one recipient address will be logged also.

@@

    \item [DELETE] Deals with mail deleted by the administrator using the
        administrative command \daemon{postsuper}.  This action adds a
        dummy recipient address if required, then invokes the COMMIT action
        to handle adding the mail to the database.  The complication this
        action deals with is described fully in \sectionref{Mail deleted
        before delivery is attempted}.

@@

\subsubsection{Mail deleted before delivery is attempted}

\label{Mail deleted before delivery is attempted}

Postfix logs the recipient address when delivery of a mail is attempted, so
if delivery has yet to be attempted the parser cannot determine the
recipient address or addresses.  This is a problem when mail is arriving
faster than Postfix can attempt delivery, and the administrator deletes
some of the mail (because it's the result of a mail loop, mail bomb, or
joe-job) before Postfix has had a chance to try to deliver it.  In this
case the recipient address will not have been logged, so a dummy recipient
address needs to be added, as every mail is required to have at least one
recipient.  This complication is generally observed in few log files, but
typically when it does occur there will be many instances of it, closely
grouped.
